Advanced Reactor Fuel Cycles Molten Salt Reactor Design

### 2016-10-19 MSR-design.tex
Add MSRE fuel salt, coolant salt, and graphite operating temperatures.

\title{Advanced Reactor Fuel Cycle (ARFC) Molten Salt Reactor Design}
\author{Alexander Lindsay, Andrei Rykhlevskii, Kathryn Huff} % As you author additions/changes, add
                                      % your name here!!!

@@

@@

\subsubsection{Reactor vessel}

A cutaway diagram of the \gls{MSRE} reactor is shown in \cref{fig:MSRE_reactor}. Fuel
enters through the inlet and is distributed evenly along the outside of the
reactor vessel in a small annular region of 1 inch
width \cite{robertson_msre_1965}. The fluid fuel then flows turbulently 
downward, cooling the reactor vessel to within 5\textdegree F of the entering 
bulk salt temperature (1175 \textdegree F) \cite{robertson_msre_1965}. This 
down-flowing region is typically referred to as the \textit{downcomer}. The 
fuel salt is collected in the lower plenum (lower \textit{space}) of the 
reactor where straightening vanes remove the salt's rotational motion. The salt 
then flows upward through approximately 1140 0.4-inch by 1.2-inch fuel channels 
that are formed by machining grooves in the sides of the graphite moderator 
columns \cite{robertson_msre_1965}; these are the regions where fission occurs. 
The graphite columns themselves have 2-inch by 2-inch square cross sections. 
Nominal core volume is 90 ft$^3$, with 70 ft$^3$ occupied by graphite and 20 
ft$^3$ occupied by fuel \cite{robertson_msre_1965}.
After flowing through the core channels, the fuel collects in the upper plenum
before exiting through the outlet.

@@

@@

@@

with the last term representing the effect of fuel advection. In
\cite{kophazi_development_2009}, \cref{eq:neutrons} is solved only in the
reactor core, whereas \cref{eq:precursors} is solved both in the core and in the
\gls{MSRE} external primary circuit. The external loop is modeled as a single pipe
with a volume equal to the total volume of all the external components and
pipelines of the primary circuitry \cite{kophazi_development_2009}. It is not
entirely clear how pipe cross sectional areas are handled, which would determine
the transit times of individual pieces of fluid and subsequently the
distribution of delayed neutron production in the external primary circuit. That
is a piece that would likely need to be included in detailed ARFC accident
models.

@@

\subsection{ARFC \gls{MSRE} Model Design Plan}

@@

\FloatBarrier

\bibliographystyle{unsrt}
\bibliography{MSR-design}
\end{document}

>>>>

\title{Advanced Reactor Fuel Cycles Molten Salt Reactor Design}
\author{Alexander Lindsay, Andrei Rykhlevskii, Kathryn Huff} % As you author additions/changes, add
                                      % your name here!!!

@@

The melting point of FLiBe is 459 \textdegree C and the boiling point is 1430
\textdegree C \cite{_flibe_2016}.

@@

\subsubsection{Reactor vessel}
\label{sec:reactor}

A cutaway diagram of the \gls{MSRE} reactor is shown in
\cref{fig:MSRE_reactor}. Fuel enters through the inlet and is distributed evenly
along the outside of the reactor vessel in a small annular region of 1 inch
width \cite{robertson_msre_1965}. The fluid fuel then flows turbulently
downward, cooling the reactor vessel to within 5\textdegree F of the entering
bulk salt temperature (1175 \textdegree F) \cite{robertson_msre_1965}. This
down-flowing region is typically referred to as the \textit{downcomer}. The fuel
salt is collected in the lower plenum (lower \textit{space}) of the reactor
where straightening vanes remove the salt's rotational motion. The salt then
flows upward through approximately 1140 0.4-inch by 1.2-inch fuel channels that
are formed by machining grooves in the sides of the graphite moderator columns
\cite{robertson_msre_1965}; these are the regions where fission occurs.  The
graphite columns themselves have 2-inch by 2-inch square cross sections.
Nominal core volume is 90 ft$^3$, with 70 ft$^3$ occupied by graphite and 20
ft$^3$ occupied by fuel \cite{robertson_msre_1965}.  After flowing through the
core channels, the fuel collects in the upper plenum before exiting through the
outlet. At 10 MW, 0.6 MW of heat is generated in the graphite, 1.4 MW heat is
generated in the fuel outside the nominal core, and 8.0 MW is generated in the
fuel within the core \cite{robertson_msre_1965}. I'm very curious how those
numbers were determined.

@@

Flow in fuel channels is laminar; because both the graphite and fueld have good
thermal conductivities, the maximum temperature of the graphite is only about 60
\textdegree F above the mixed mean temperature of the adjacent fuel. Average and
maximum temperatures of graphite are 1255 \textdegree F and 1300 \textdegree F
respectively. The temperature of the fuel leaving the hottest channel in the
core is about 1260 \textdegree F. The fuel leaves the top of the reactor at 7
psig and an average temperature of 1225 \textdegree F \cite{robertson_msre_1965}.

@@

\subsubsection{Heat Exchanger}
\label{sec:heatx}

As mentioned in \cref{sec:reactor}, the fuel salt enters the reactor at 1175
\textdegree F and leaves at 1225 \textdegree F. In order to cool the fuel salt
back down to its reactor inlet temperature, the coolant salt flows through the
heat exchanger at 850 \gls{gpm} with inlet and outlet temperatures of 1025
\textdegree F and 1100 \textdegree F respectively \cite{robertson_msre_1965}.

@@

with the last term representing the effect of fuel advection. In
\cite{kophazi_development_2009}, \cref{eq:neutrons} is solved only in the
reactor core, whereas \cref{eq:precursors} is solved both in the core and in the
\gls{MSRE} external primary circuit. The external loop is modeled as a single pipe
with a volume equal to the total volume of all the external components and
pipelines of the primary circuitry \cite{kophazi_development_2009}. It is not
entirely clear how pipe cross sectional areas are handled, which would determine
the transit times of individual pieces of fluid and subsequently the
distribution of delayed neutron production in the external primary circuit. That
is a piece that would likely need to be included in detailed \gls{ARFC} accident
models.

@@

\subsection{\gls{ARFC} \gls{MSRE} Model Design Plan}

@@

\FloatBarrier
\clearpage
\printglossary[type=\acronymtype]
\bibliographystyle{unsrt}
\bibliography{MSR-design}
\end{document}
